---
title: "Simulating pairwise social interactions datasets"
author: "R.E. Wijnhorst"
date: "2023-09-14"
---

```{r setup, include=FALSE}
require(MASS)
```

Step 1. Enter study design dimensions and the number of iterated datasets for balanced datasets of pairwise interactions

```{r sampling dimensions and number of datasets}
n_ind     = 200 # number of individuals 
partners  = 4   # number of partners per individual
repeats   = 1   # number of repeats of unique dyadic interaction
iterations= 100 # number of data sets
```

Step 2. Enter effect sizes. 

In order to simulate data with only a DGE and IGE (for variance-partitioning approaches), set Mpsi=0 and Vpsi=0. 
For trait-based or hybrid data without variation in responsiveness, set Vpsi=0. 
Optional: simulate measurement error on the opponent trait with Vxe. 

```{r parameters}
# Means
Malpha   = 1      # mean focal effect
Mepsilon = 0      # mean partner effect
Mpsi     = 0.3    # interaction coefficient psi
Mx       = 0      # mean-centered covariate/opponent trait

# Variances
Valpha   = 0.2    # variance direct effect (mean behaviour)
Vepsilon = 0.01   # variance social partner effect (residual impact)
Vpsi     = 0.1    # variance in slopes (social responsiveness)
Vx       = 1      # variance social partner phenotype (impact covariate)
Ve       = 0.6    # residual variance
Vxe      = 0.1    # measurement error or variation of labile trait

# Correlations between variance components
r_alpha_epsilon  =  0       # part of Cov int-phi
r_alpha_psi      =  -0.6    # part of Cov int-psi 
r_epsilon_psi    =  -0.6    # part of Cov psi-phi 
r_alpha_x        =  0.6     # part of Cov int-phi 
r_psi_x          =  -0.6    # part of Cov psi-phi 
r_epsilon_x      =  0
```

Step 3. Run code to simulate datasets. Dataframes are bundled in lists and saved as ".RDS" in the format that can be readily analysed with rstan.

```{r function sim}
# Random balanced design function
sampling_design_balanced <- function(n_ind,partners,repeats){
ind_seq <- 1:n_ind 
IDi <- rep(ind_seq, partners) 
IDj = vector("list", length = n_ind*partners) 
  for (i in 1:partners) {
    seq_j <- c(ind_seq[-c(1:i)], ind_seq[1:i])
    IDj[[i]] <- seq_j   }
IDj = do.call(c, IDj)
    df <- data.frame(IDi,IDj)
    df <- df[rep(seq_len(nrow(df)), repeats),] 
    df
}

# Means & Variance G-matrix for multivariate normal distribution
M<-c(Malpha,Mepsilon,Mpsi,Mx)
G<-matrix(NA,4,4)

G[1,1]<-Valpha
G[2,2]<-Vepsilon
G[3,3]<-Vpsi
G[4,4]<-Vx

G[1,2]<-G[2,1]<-r_alpha_epsilon*(sqrt(Valpha*Vepsilon)) 
G[1,3]<-G[3,1]<-r_alpha_psi*(sqrt(Valpha*Vpsi)) 
G[2,3]<-G[3,2]<-r_epsilon_psi*(sqrt(Vepsilon*Vpsi))
G[4,1]<-G[1,4]<-r_alpha_x*(sqrt(Valpha*Vx))
G[4,2]<-G[2,4]<-r_epsilon_x*(sqrt(Vepsilon*Vx))
G[4,3]<-G[3,4]<-r_psi_x*(sqrt(Vpsi*Vx))

# Create phenotypic traits of individuals and outcome of interactions (requires study design)
sim<-function(df, M, G, Ve) {
a=as.data.frame(mvrnorm(n_ind, M, G))
colnames(a)<-c("alpha", "epsilon", "psi", "x")
a$ID=1:n_ind

df$alpha_i=a[match(df$IDi,a$ID),"alpha"]
df$epsilon_j=a[match(df$IDj,a$ID),"epsilon"]
df$psi_i=a[match(df$IDi,a$ID),"psi"]
df$xj=a[match(df$IDj,a$ID),"x"]
df$x_ijk <- df$xj + rnorm(nrow(df),0, sqrt(Vxe))
df$e_ijk<-rnorm(nrow(df),0, sqrt(Ve))

df$z_i=df$alpha_i + df$psi_i*df$x_ijk + df$epsilon_j + df$e_ijk # Phenotypic Equation
df
}

# Simulate datasets and create lists
df<-sampling_design_balanced(n_ind, partners, repeats) # see Random balanced design function
dfl <- list()
for(i in 1:iterations){
df2<-sim(df, M, G, Ve)
xi<-aggregate(df2$x_ijk, list(IDi = df2$IDj), mean)
dfl[[i]]<-list(n_obs = nrow(df2),
                 n_ind=length(unique((df2$IDi))),
                 individual=df2$IDi,
                 opponent=df2$IDj,
                 xj=df2$x_ijk,  # model input is the observed xj
                 xi=xi$x,
                 z=df2$z_i)
}

# List saving: dfl_N(total observations)_N(indidviduals)_N(social partners)_N(repeats)
saveRDS(dfl, paste0('dfl_',n_ind*partners*repeats,sep="_",n_ind,sep="_",partners,sep="_",repeats,"x.RDS"))
```